---
title: "Practical Machine Learning Project"
author: "Eric Allen"
date: "5/17/2019"
output: html_document
---

## Study Objective

The goal of the project is to predict the manner in which exercises were performed. This is the "classe" variable in the training set (A = correctly, B-E = different variations of incorrectly). In this project, we will use data from sensors on the belt, forearm, arm, and dumbbell of 6 participants.  

## Loading the training data

```{r}

file <- "pml-training.csv"
if (!file.exists(file))
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",file)
df <- read.csv("pml-training.csv")
dim(df)

```

## Cleaning the data 

Initial data exploration uncovered fields with data that did not come from the sensors (see appendix for field name listing), fields with nearly all missing values and fields with near zero variance.  These fields were removed from the data prior to training the model. 

```{r}

# remove non-sensor data
dfc <- df[,c(8:160)]

# testing for columns with NA
nacnt <- colSums(is.na(dfc))
table(nacnt) # 86 columns have no missing values where 67 have 19,216 out of 19,622 missing
# remove fields with NA in them
dfc <- dfc[,nacnt==0]

# remove near zero variance variables

library(caret)
NZdf <- nearZeroVar(dfc,saveMetrics = TRUE)
NZdf$rownum <- 1:nrow(NZdf)
nzvIndex <- NZdf[NZdf$nzv==FALSE,]$rownum
dfc <- dfc[,!NZdf$nzv]

```

## Partition the data for validation 

The training data was partitioned into a training and test data sets for validation purposes.  

```{r}
# create training partitions for validation

set.seed(3825)
inTrain <- createDataPartition(y=dfc$classe, p=0.75, list=FALSE)
dftrn <- dfc[inTrain,]
dft <- dfc[-inTrain,]
dim(dftrn)
dim(dft)
```

## Building the Model

The model was built using Random Forest with cross validation. 

```{r cache=TRUE}

library(doMC)
registerDoMC(cores=3)

set.seed(3825)

control <- trainControl(method="cv", number=10, allowParallel = TRUE)
modRF <- train(classe ~ ., method = "rf", data = dftrn, trControl=control)
plot(modRF,main="Accuracy by Predictor Count")
cvconfm <- confusionMatrix.train(modRF)
cvconfm

```

10-fold cross-validation shows that `r modRF$results[modRF$results[2]==max(modRF$results[2]),][[1]]` randomly selected predicors produces the highest accuracy of `r modRF$results[modRF$results[2]==max(modRF$results[2]),][[2]]` with an accuracy SD of `r modRF$results[modRF$results[2]==max(modRF$results[2]),][[4]]`. The mean estimated out of sample Accuracy is `r sum(diag(cvconfm$table))/100`. 

```{r}
confm <- confusionMatrix(dft$classe,predict(modRF,newdata = dft))
```

Using the model to predict test partition shows and Accuracy of `r confm$overall[[1]]` with a confidence interval from `r confm$overall[[3]]` to `r confm$overall[[4]]`.

```{r}
confm
```

## Quiz

```{r}
# download and prep quiz data

file <- "pml-testing.csv"
if (!file.exists(file))
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",file)
dftesting <- read.csv("pml-testing.csv")
dim(dftesting)

# select only the variables used
dfval <- dftesting[,names(dftrn[,-53])]
dim(dfval)

quiz <- predict(modRF,dfval)
answers <- as.data.frame(quiz)

```

## References
More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Appendix
```{r}
names(df)
```

